Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-02-20 ms/SI.rmd
Update MS

```{r ancova}
# Format table
ancova_table <-
  tribble(
  ~variable, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "temp_sd", "SD temp.",
  "vpd_min", "Min. VPD",
  "vpd_mean", "Mean VPD",
  "vpd_sd", "SD VPD",
) %>%
  left_join(climate_ancova_results, by = "variable") %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(Response, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)

@@

phylosig_binary_strict_footer <- "'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."

@@

'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.

@@

corr_evo_strict_footer <- "LL, log likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Lower log likelihood for the dependent model indicates correlation with growth habit."

@@

```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
cwsd_scatterplot
```

**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`. Epiphytic communities in green, terrestrial communities in brown. Points randomly jittered along x-axis to reduce overlap. Boxplots show median values (bold lines) by growth habit. Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range. Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size. Response variable abbreviations: SLA, specific leaf area.

@@

```{r Fig-S2, fig.height = 8, fig.width = 4, fig.cap = ""}
func_div_corr_plot
```

@@

>>>>

```{r ancova}
# Format table
ancova_table <-
  tribble(
  ~variable, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "temp_sd", "SD temp.",
  "vpd_min", "Min. VPD",
  "vpd_mean", "Mean VPD",
  "vpd_sd", "SD VPD",
) %>%
  left_join(climate_ancova_results, by = "variable") %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit x elevation",
      TRUE ~ term
    )
  ) %>%
  select(Response, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)

@@

phylosig_binary_strict_footer <- "D, Fritz and Purvis' D (Fritz & Purvis, 2010); 'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."

@@

\D, Fritz and Purvis' \D (Fritz & Purvis, 2010); 'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.

@@

corr_evo_strict_footer <- "LL, log likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Higher (less negative) log likelihood for the dependent model indicates correlation with growth habit."

@@

```{r corr-plot, fig.height = 8, fig.width = 4, fig.cap = ""}
func_div_corr_plot
```

@@

```{r cwsd, fig.height = 6, fig.width = 7, fig.cap = ""}
cwsd_scatterplot
```

**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`. Epiphytic communities in green, terrestrial communities in brown. Points randomly jittered along x-axis to reduce overlap. Boxplots show median values (bold lines) by growth habit. Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range. Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001. **(a)**, traits related to plant size **(b)**, traits not related to plant size. Response variable abbreviations: SLA, specific leaf area.

@@

**Fritz SA, Purvis A**. **2010**. Selectivity in mammalian extinction risk and 
threat types: A new measure of phylogenetic signal strength in binary traits. 
*Conservation Biology* **24**: 1042--1051.

### 2020-02-20 ms/manuscript.Rmd
Make join explicit
Update MS

We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions. We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] for each model and retained models within 2 AIC~c~ of the lowest (\ie, 2 delta AIC~c~). From these selected we the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
Mt. Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J. Nitta, pers. obs.).

@@

We used distance-based methods to analyze community phylogenetic and functional diversity.
Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community [MNTD; @Webb2000; @Webb2002].
Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
For phylogenetic distances, we used the tree of @Nitta2017.
For functional distances, we log-transformed and scaled traits as in the PCA, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coefficient [@Gower1971] with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (Supporting Information).

@@

We tested the effects of climate and growth habit on diversity metrics using full-subsets multiple regression with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2()` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
Therefore, we also constructed linear models for each diversity metric in response to elevation and growth habit including all sites using the same model selection procedure as in the climate analysis.
We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013]. 
In the case the spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the \package{spAMM} R package [@Rousset2014].

@@

Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (standard or phylogenetic PCA) (`r figure("pca")`**c**, **d**).
The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
In the standard PCA, size-related traits tended to load on PC1, with SLA, number of pinnae pairs, and degree of dissection on PC2.
Loadings in the phylogenetic PCA are similar, but size-related traits are approximately reversed relative to the standard PCA.

@@

Most sporophyte traits, particularly those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r table("pics")`).
Epiphytic species tend to have smaller values for all traits (`r table("pics")`; contrasts were made by subtracting epiphytic from terrestrial values).
Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
For the gametophyte traits, only morphotype (non-cordate vs. cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r table("corr-evo")`).
Similar results were obtained using the 'strict' dataset (`r s_table("corr-evo-strict")`).

@@

# Make tibble of abundances
abundances <-
  comm %>%
  mutate(abundance = 
           comm %>%
           select_if(is.numeric) %>%
           rowSums()
  ) %>%
  select(species, abundance)

# Join this with trait data
traits_with_abun <- left_join(
  fern_traits, abundances
)

@@

Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }= `r get_mod_result(model_summ_print, "dissection", "p")`).
Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).

We observed a total of `r traits_with_abun %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
Species richness increased with elevation for both epiphytes and terrestrial communities (`r figure("comm-div")`).
Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r figure("comm-div")`).
No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).
Spatial autocorrelation was not detected in any of the linear models for the community diversity metrics (`r s_table("linear-fit")`).

@@

Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r figure("cwm-div")`, `r table("pics")`).
Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs. slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; @Diaz2015].
We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes when taking elevation into account (`r figure("cwm-div")`).
This may reflect adaptations to specific niches that are not captured by a binary categorization of growth habit.
For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Humata anderssonii* (Mett.) C.Chr. (a twig epiphyte), *Acrostichum aureum* L. (a terrestrial mangrove specialist), and *Paesia divaticatissima* (Dryand.) Copel. (a terrestrial species occurring on high-elevation ridges) (`r s_table("trait-values")`).
Each of these species is likely to experience prolonged periods of drought-stress. 
In-depth characterization of such microhabitats was beyond the scope of our study and should be investigated further in the future.

@@

@@

**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}{\meter} on Moorea, French Polynesia. Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`. Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD). 'Epiphytic' dataloggers were placed at \circa{ }\SI{2}{\meter} on trees, and 'terrestrial' dataloggers placed at ground level. Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown. Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05. Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05. Partially transparent colors indicate outliers excluded from the models (see Methods).

@@

**`r figure("cwm-div")`** Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia. Response variable abbreviations: SLA, specific leaf area. Epiphytic communities in green, terrestrial communities in brown. Error bars are each one standard deviation (values below zero not shown). Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods). \tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01. Results for frond width were very similar to frond length and not shown.

@@

# FIXME: need to manually add "Sporophyte traits" and "Gametophyte traits" to
# table in docx (this happens with gt or kable), and fix italics, superscript, etc.
read_csv(here::here("data/table_1.csv")) %>%
  rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance, `Reference(s)` = reference) %>%
  column_to_rownames("trait") %>%
  knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE) %>%
  kable_styling(full_width = FALSE) %>%
  pack_rows("Sporophyte traits", 1, 7) %>%
  pack_rows("Gametophyte traits", 8, 11) %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "12em") %>%
  column_spec(4, width = "8em")
```

@@

LogL, log-likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial growth). Traits with a significantly lower log likelihood for the dependent model are significantly correlated with growth habit.

>>>>

We calculated daily maximum, mean, minimum, and standard variation of temperature, RH, and VPD for each datalogger, then overall means from the daily values.
We analyzed correlation between the mean values with Pearson's correlation coefficient, and retained a set with correlation coefficients less than 0.9 (a single variable was arbitrarily selected from each set of correlated variables; hereafter, 'climatic variables').
Variables relating to VPD, which have no meaning if less than zero, and some other variables that had skewed distributions, were square-root transformed prior to fitting models.
We tested for differences between epiphytic and terrestrial dataloggers while accounting for elevation with ANCOVA, using growth habit as the categorical variable and elevation as the covariate.
We identified the best-fitting linear model for each climatic variable in response to growth habit and elevation by constructing a set of models including the effect of each independent variable alone, in combination, and in combination with their interactions. We then calculated the corrected Akaike Information Criterion [AIC~c~; @Akaike1973] for each model and retained models within 2 AIC~c~ of the lowest (\ie, 2 delta AIC~c~). From these, we selected the model with the combination of the fewest non-significant parameters and lowest AIC~c~ (for VPD, we the selected the model producing the smallest amount of negative predicted values).
A single outlier plot with unusually high temperature and low humidity (Mt. Rotui, 830 m plot) was excluded from ANCOVA and linear models.
Mt. Rotui is an isolated peak with greater exposure than the other two mountains where the rest of the plots were established (J. Nitta, pers. obs.).

@@

We used distance-based methods to analyze community phylogenetic and functional diversity.
Mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) describe the mean distance between all species in a community (MPD) or the mean distance between the closest pairs of species in a community [MNTD; @Webb2000; @Webb2002].
Although these metrics were originally formulated for phylogenetic distances, they can be applied in the same way to trait-based distances, enabling a direct comparison between the two [@Sessa2018].
For phylogenetic distances, we used the tree of @Nitta2017.
For functional distances, we log-transformed and scaled traits as in the PCA, excluded frond width and length (both frond size traits highly correlated with stipe length), then calculated the distance matrix using Gower's dissimilarity coefficient [@Gower1971] with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011], with equal weighting for sporophyte and gametophyte traits.
To test for the effect of different trait selection schemes, we compared the results of running the same analyses for various sets of traits with and without log-transformation (`r s_figure("corr-plot")`).

@@

We tested the effects of climate and growth habit on diversity metrics using full-subsets multiple regression with general additive models (GAMs) as implemented in the \package{FSSgamm} R package [@Fisher2018].
For each diversity metric, we built a set of models (Gaussian family, with plot as a random effect) taking into account all combinations of explanatory variables and their pairwise interactions, and ranked these by AIC~c~ to find the models that best explain the data.
Our dataset included `r (nrow(div_metrics_all) - nrow(climate_select)) %>% english2()` communities that lacked climate data and could not be included in full-subsets multiple regression analysis.
Therefore, we also constructed linear models for each diversity metric in response to elevation and growth habit including all sites using the same model selection procedure as in the climate analysis.
We checked for spatial autocorrelation in the model residuals with Moran's *I* [@Moran1950] as implemented in the \package{spdep} R package [@Bivand2013]. 
If spatial autocorrelation was detected, a generalized linear mixed model was fit with a Mátern spatial correlation structure using the \package{spAMM} R package [@Rousset2014].

@@

Terrestrial and epiphytic species are partly differentiated in trait space, occupying mostly distinct areas but overlapping in the middle regardless of method used (`r figure("pca")`**c**, **d**).
The first two PC axes explained `r percent(pca_std_cum_var)` and `r percent(pca_phy_cum_var)` of the variation in the trait data when analyzed using standard and phylogenetic PCA, respectively.
In the standard PCA, size-related traits tended to load on PC1, with SLA, number of pinnae pairs, and degree of dissection on PC2.
Loadings in the phylogenetic PCA are similar, but size-related traits are approximately reversed relative to the standard PCA.

@@

Most sporophyte traits, particularly those related to plant size, are significantly correlated with growth habit while taking phylogeny into account (`r table("pics")`).
Epiphytic species tend to have smaller values for all traits (`r table("pics")`).
Short stipe length in particular appears to be strongly correlated with epiphytic growth (phylogenetically independent contrasts, \pval{ }`r stipe_pval`), as well as frond width (\pval{ }`r width_pval`) and degree of laminae dissection (\pval{ }`r dissection_pval`).
For the gametophyte traits, only morphotype (non-cordate vs. cordate) was significantly correlated with growth habit (Pagel's test of correlated evolution, \pval{ }`r morphotype_pval`; `r table("corr-evo")`).
Similar results were obtained using the 'strict' dataset (`r s_table("corr-evo-strict")`).

@@

# Make tibble of abundances
abundances <-
  comm %>%
  pivot_longer(cols = -species, values_to = "abundance", names_to = "site") %>%
  assert(not_na, abundance) %>%
  group_by(species) %>%
  summarize(abundance = sum(abundance)) %>%
  ungroup()

# Join this with trait data
traits_with_abun <- left_join(
  fern_traits, abundances, by = "species"
)

@@

Community-weighted mean values (CWMs) were smaller for epiphytic than terrestrial ferns for size-related traits (frond length, frond width, stipe length, and rhizome diameter) and number of pinnae pairs (\tval-test, all \pval{ }< 0.05; `r figure("cwm-div")`).
Spatial autocorrelation was detected in the linear model for SLA (`r s_table("linear-fit")`), so a spatial generalized mixed model was used instead (see Methods). 
Specific leaf area decreased with elevation, and was slightly lower for epiphytic relative to terrestrial communities (spatial generalized linear mixed model, nu = `r sla_nu`, rho = `r sla_rho`, AIC = `r sla_AIC`; `r s_table("spatial-sla")`).
Degree of lamina dissection decreased with elevation in epiphytes, but increased in terrestrial communities (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "dissection", "r")`, \pval{ }`r get_mod_result(model_summ_print, "dissection", "p")`).
Frond length and width and rhizome diameter decreased slightly with elevation for terrestrial communities, while increasing slightly for epiphytic communities (`r figure("cwm-div")`).

We observed a total of `r abundances %>% filter(abundance > 0) %>% pull(species) %>% n_distinct` \species across all plots, including `r n_epi` epiphytic and `r n_ter` terrestrial spp.
Mean species richness was `r mean_rich %>% filter(habit == "epiphytic") %>% pull(stat)` \species per plot for epiphytes and `r mean_rich %>% filter(habit == "terrestrial") %>% pull(stat)` \species per plot for terrestrial species (not significantly different; two-sided \tval-test, \pval{ }= `r get_mod_result(t_test_print, "ntaxa", "p")`; `r figure("comm-div")`).
Species richness increased with elevation for both epiphytes and terrestrial communities (`r figure("comm-div")`).
Epiphytic communities were more diverse than terrestrial ones as measured by both MPD~func~ and MNTD~func~ (two-sided \tval-test, `r both`\pval{ }`r p_mpd_func_mntd_func`).
Functional diversity increased along the gradient for both terrestrial and epiphytic communities when measured with MPD~func~ (linear model, \rval^2^ = `r get_mod_result(model_summ_print, "mpd.obs.z.func", "r")`, \pval{ }`r get_mod_result(model_summ_print, "mpd.obs.z.func", "p")`), but not MNTD~func~.
Epiphytic communities were more phylogenetically clustered than terrestrial communities overall when measured using MNTD~phy~ (two-sided \tval-test, \pval{ }`r get_mod_result(t_test_print, "mntd.obs.z.phy", "p")`) but not MPD~phy~ (\pval{ }= `r get_mod_result(t_test_print, "mpd.obs.z.phy", "p")`; `r figure("comm-div")`).
No significant trend was detected in phylogenetic community structure with elevation in either terrestrial or epiphytic communities (`r figure("comm-div")`).
Spatial autocorrelation was not detected in any of the linear models for the community diversity metrics (`r s_table("linear-fit")`).

@@

Another trait that defied simple classification into epiphytic or terrestrial groups was specific leaf area (SLA; `r figure("cwm-div")`, `r table("pics")`).
Specific leaf area is part of the leaf economic spectrum (LES), which posits that leaf traits including SLA, N and C content, and leaf life span vary along well-defined axes such that plants generally fall into one of two categories: fast-growing species with high SLA, low nutrient content, and short life spans vs. slow-growing species with low SLA, high nutrient content, and long life spans [@Wright2004a; @Diaz2015].
We expected that epiphytes would exhibit lower SLA, given their more extreme growing conditions.
In contrast, we observed overlapping, wide ranging SLA values in both terrestrial and epiphytic communities, with only slightly lower values in epiphytes when taking elevation into account (`r figure("cwm-div")`).
This may reflect adaptations to specific niches that are not captured by a binary categorization of growth habit.
For example, the following species ranked among those with the five lowest SLA values (\ie, thick, well-protected leaves) on Moorea: *Davallia anderssonii* Mett. (a twig epiphyte), *Acrostichum aureum* L. (a terrestrial mangrove specialist), and *Paesia divaticatissima* (Dryand.) Copel. (a terrestrial species occurring on high-elevation ridges) (`r s_table("trait-values")`).
Each of these species is likely to experience prolonged periods of drought-stress. 
In-depth characterization of such microhabitats was beyond the scope of our study and should be investigated further in the future.

@@

**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")`

@@

**`r figure("climate")`** Selected microclimatic variables of study plots along an elevational gradient from \SI{200}{\meter} to \SI{1200}{\meter} on Moorea, French Polynesia. Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`. Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD). 'Epiphytic' dataloggers were placed at \circa{ }\SI{2}{\meter} on trees, and 'terrestrial' dataloggers placed at ground level. Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown. Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval{ }< 0.05. Asterisks indicate statistical significance of linear model; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01. Partially transparent colors indicate outliers excluded from the models (see Methods).

@@

**`r figure("cwm-div")`** Change in community-weighted mean (CWM) values of epiphytic and terrestrial fern communities along an elevational gradient on Moorea, French Polynesia. Response variable abbreviations: SLA, specific leaf area. Epiphytic communities in green, terrestrial communities in brown. Error bars are each one standard deviation (values below zero not shown). Trendlines and \rval^2^ shown for significant relationships with elevation or the interaction of elevation and growth habit as determined by a linear model at \pval{ }< 0.05 (\rval^2^ not shown for SLA, which was fit with a spatial generalized linear mixed model; see Methods). \tval shown for significant differences between means of CWMs by growth habit as determined by a two-sided \tval-test at \pval{ }< 0.05; \*\*\* = \pval{ }< 0.001. Results for frond width were very similar to frond length and not shown.

@@

# FIXME: need to manually add "Sporophyte traits" and "Gametophyte traits" to
# table in docx (this happens with gt or kable), and fix italics, superscript, etc.
read_csv(here::here("data/table_1.csv"), col_types = "cccc") %>%
  rename(`Data type (unit)` = data_type, `Functional significance` = ecological_significance, `Reference(s)` = reference) %>%
  column_to_rownames("trait") %>%
  knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE) %>%
  kable_styling(full_width = FALSE) %>%
  pack_rows("Sporophyte traits", 1, 7) %>%
  pack_rows("Gametophyte traits", 8, 11) %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "12em") %>%
  column_spec(4, width = "8em")
```

@@

LogL, log-likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial growth). Traits with a significantly higher (\ie, less negative) log likelihood for the dependent model are correlated with growth habit.
